IMPC Mouse data - Variance in sex differences

### 2019-07-05 scripts/2019_4June_IMPC variance and sex difference.Rmd
Added a new Shiny app to explore meta-analytic results. Has a theme selector built in so you can check out what works and what doesn't. I can hardwire that later when people pick one and agree. I also looked through code. Fixed some small types, modified results.grouping table to include traits and added export of this table so it can be used for calculations of percent diff in shiny apps

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne & Felix Zajitschek"
date: "2nd April 2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
---

@@

# Loading functions that are necessary
### Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a the cleaned up file which is less computing intensive. However, cvs
```{r}

@@

  mydata[i,] %>% 
  select(strain_name, production_center) %>%
    mutate(
      effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can ve loaded and cleaned using the da cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than th e .csv - which we can still provide for those who absolutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")

@@

```{r}
results.alltraits.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*13), ncol = 13))) #number of individual results per trait = 10
names(results.alltraits.grouping) <- c("id", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se" , "sampleSize")

## LOOP, to run meta-analysis on all traits
The loop combines the functions mentioned above and fills the data matrix with results from our meta analysis. 
Error messages indicate traits that either did not reach convergence, or that did not return meaningful results in the meta-analysis, due to absence of variance. Those traits will be removed ion later steps, outlined below.

@@

#lnCVR,  log repsonse-ratio of the coefficient of variance    
    cvr <- metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(~1| strain_name, ~1|production_center, results.alltraits.grouping[t, 2] <- cvr$b
    results.alltraits.grouping[t, 3] <- cvr$ci.lb
    results.alltraits.grouping[t, 4] <- cvr$ci.ub
    results.alltraits.grouping[t, 5] <- cvr$se

@@

cv <- metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(~1| strain_name, ~1|production_center,                                                             ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000),  verbose=F, data = results)
    results.alltraits.grouping[t, 6] <- cv$b
    results.alltraits.grouping[t, 7] <- cv$ci.lb
    results.alltraits.grouping[t, 8] <- cv$ci.ub
    results.alltraits.grouping[t, 9] <- cv$se

@@

        results.alltraits.grouping[t, 14] <- means$k

@@

>>>>

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne & Felix Zajitschek"
date: "2nd April 2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
runtime: shiny
---

@@

# Loading functions that are necessary
### Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a cleaned up file which is less computing intensive. However, cvs
```{r}

@@

  mydata[i,] %>% 
  select(strain_name, production_center, trait) %>%
    mutate(
      effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than the .csv - which we can still provide for those who absolutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")

@@

```{r}
results.alltraits.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*14), ncol = 14))) #number of individual results per trait = 10
names(results.alltraits.grouping) <- c("id", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se" , "sampleSize", "trait")
```

## LOOP, to run meta-analysis on all traits
The loop combines the functions mentioned above and fills the data matrix with results from our meta analysis. 
Error messages indicate traits that either did not reach convergence, or that did not return meaningful results in the meta-analysis, due to absence of variance. Those traits will be removed in later steps, outlined below.

@@

#lnCVR,  log repsonse-ratio of the coefficient of variance    
    cvr <- metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(~1| strain_name, ~1|production_center, ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000),  verbose=F, data = results)

    results.alltraits.grouping[t, 2] <- cvr$b
    results.alltraits.grouping[t, 3] <- cvr$ci.lb
    results.alltraits.grouping[t, 4] <- cvr$ci.ub
    results.alltraits.grouping[t, 5] <- cvr$se

@@

cv <- metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(~1| strain_name, ~1|production_center,~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000),  verbose=F, data = results)

    results.alltraits.grouping[t, 6] <- cv$b
    results.alltraits.grouping[t, 7] <- cv$ci.lb
    results.alltraits.grouping[t, 8] <- cv$ci.ub
    results.alltraits.grouping[t, 9] <- cv$se

@@

        results.alltraits.grouping[t, 14] <- means$k
        results.alltraits.grouping[t, 15] <- unique(results$trait)

@@

# Now that we have a "results" table with each of the meta-analytic means for all effect sizes of interest, we can use this table as part of the Shint App, which will then be able to back calculate the percentage differences between males and females for mean, variance and coefficient of variance. We'll export and use this in the Shiny App. **Note that I have not dealt with convergence issues in some of these models, and so, this will need to be done down the road**

trait_meta_results <- write.csv(results.alltraits.grouping, file = "export/trait_meta_results.csv")
